---
title: "03_random_forrest_spore"
output: html_document
date: "2023-04-06"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(magrittr)
library(tidymodels)
library(cowplot)
library(modelr)
library(naniar)
library(patchwork)
```

## Reading in data

```{r}
D <- readRDS("./data/bacdive_growth_ribdif.rds")
# Pathogen, set nan to 0
Dt_tmp1 <- D %>% 
  mutate(pathogen_any = ifelse(is.na(pathogen_any), "P0", pathogen_any))

# Set nan to neutrophile
Dt_tmp2 <- Dt_tmp1 %>% 
  mutate(PH.range = ifelse(is.na(PH.range), "Pneutrophile", PH.range))

# Get pseudogenes as a percent
Dt_tmp3 <- Dt_tmp2 %>% 
  mutate(pseudogenes_percent = 100*(pseudogenes/total_genes))

# Selecting data and filtering in phylums
D_all <- Dt_tmp3 %>%  
  filter(phylum %in% c("Bacillota","Actinomycetota")) %>% 
  select(species,last_col(42):last_col())  %>% 
  mutate(growth_temp = as.double(growth_temp)) 

# Filter out nans for sporeforming
D_filtered_tmp <- D_all %>% 
  filter(!is.na(sporeforming))


D %>% filter(class == "Deltaproteobacteria") %>% select(species,sporeforming)

Dall %>% left_join(D, by="species") %>% 
  filter(class.x =="Deltaproteobacteria") %>% 
  select(species,sporeforming.y,n16.x,genus.x) %>% filter(is.na(sporeforming.y))
  
Dall$n1
```

Looking at nans

```{r}
D_all %>% 
   vis_miss(cluster = T, sort_miss = T) +
    theme(
        plot.margin = margin(, 1.5, , , "cm")
  )

```

Even though growth temp and oxygen tolerance has a lot of missing values
We are still using them in the model, due to 1) they matter as seen
earlier in the pca 2) We are using them in the final larger model
anyways, so the datapoints are ging to be removed anyways In summary, we
are using the genomic data and oxygen.tol, phylum, growth_temp \###
Selecting the actual data used

```{r}
D_filtered <- D_filtered_tmp %>% 
  select(phylum, oxygen.tolerance, growth_temp,
         gc_percent, sporeforming, genes_coding, genome_components, pseudogenes_percent, ar_count) %>% 
  filter(if_all(everything(), ~!is.na(.x)))

```

### Spliting data

Using strata =, to make sure that sporeforming is equally distributed

```{r}
set.seed(07042023)
D_split <- initial_split(D_filtered, prop = 0.80, strata = sporeforming)
D_train <- training(D_split) 
D_test <- testing(D_split)
```

### Setting up model

Based on: <https://juliasilge.com/blog/sf-trees-random-tuning/> Based on
tidymodels package

```{r}
# setting up recipe
tree_rec <- recipe(sporeforming ~., data=D_train)

# Setting up model hyperparameters
tune_spec <- rand_forest(
  # number which can be sampled per tree
  mtry = tune(),
  trees = 1000,
  # min number of datapoints for a split being made 
  min_n = tune()
  ) %>%
  set_mode("classification") %>%
  set_engine("ranger")


# setting up workflow
tune_wf <- workflow() %>% 
  add_recipe(tree_rec) %>% 
  add_model(tune_spec)

```

## Tune hyperparameters

```{r}
# Set to parallel
doParallel::registerDoParallel()

# Make samples for cross-validation
trees_folds <- vfold_cv(D_train)

# Tune hyperparameters
# grid sets the 
tune_res <- tune_grid(
  tune_wf,
  # Sets the 
  resamples = trees_folds,
  grid = 20
)

```

## Picking the best model :)

### First run

```{r}
tune_res %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "AUC")
```

It seems that lower values of both the minimum numbers of data.points
per split (min_n) work better. And lower values of mtry, the number of
samples which can be choosen

Lets try to run model on more values in this interval

### Second run

```{r}
# Making grid
rf_grid <- grid_regular(
  min_n(range = c(1,15)),
  mtry(range=c(1, 8)),
  levels = 5
)


regular_res <- tune_grid(
  tune_wf,
  resamples = trees_folds,
  grid = rf_grid
)

```

Results

```{r}
regular_res %>% 
  collect_metrics()

```

AUC

```{r}
regular_res %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "AUC")
```

min_n = 4 and mtry = 2 seems the most balanced mtry -\> we need som
flexiblity in the model

finalize the model

```{r}
best_auc <- select_best(regular_res, "roc_auc")

final_rf <- finalize_model(
  tune_spec,
  best_auc
)

final_rf
```

```{r}
library(vip)

final_rf %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(factor(sporeforming) ~ . ,data=D_train) %>%
  vip(geom = "point")
```


Now lets fit final data on entire training set and evalute on testing
data

```{r}
final_wf <- workflow() %>%
  add_recipe(tree_rec) %>%
  add_model(final_rf)

final_res <- final_wf %>%
  last_fit(D_split)

final_res %>%
  collect_metrics()
```

```{r}
final_res %>% 
  collect_predictions() %>% 
  filter(.pred_class != sporeforming)

# Only 6 wrong predictions out of 59 !!  = 0.93 accurcy..
# Worse than just genus.. but appliciple 
# In all of them is it not very sure. We could change cutoff %

```

## Running model on the rest of the data

```{r}
# select entries without sporeforming info
# and select attributes
D_to_pred <- D_all %>% 
  filter(is.na(sporeforming)) %>% 
  select(-sporeforming) %>% 
  select(species,phylum, oxygen.tolerance, growth_temp, n16, 
         gc_percent, genes_coding, genome_components,  pseudogenes_percent, ar_count) %>% 
  filter(if_all(everything(), ~!is.na(.x)))
  
D_to_pred
```

Make predictions

```{r}
# Extract model
final_model <- final_res %>% 
  extract_model()

# Predict
predictions <- predict(final_model, D_to_pred, type="response")

# save in df
pred_df <- bind_cols(D_to_pred, predictions$predictions)

```

Plot best prediction confidence. Is how many trees think it. I believe

```{r}
# How is dist
pred_df %>% 
  rename(sf0 = `0`, sf1 = `1`) %>% 
  mutate(best_pred = case_when(
    sf0 > sf1 ~ sf0,
    .default = sf1
  )) %>% 
  ggplot(aes(best_pred))+ geom_histogram()

```

Looks like \~0.7 cutoff would be ok. Set those values to nan Apply
cutoff

```{r}
D_predicted <- pred_df %>% 
  rename(sf0 = `0`, sf1 = `1`) %>% 
  mutate(best_pred = case_when(
    sf0 > sf1 |sf0 > 0.7  ~ "p0",
    sf1 > sf0 | sf1 > 0.7 ~ "p1",
  )) 

```

Visualize predictions

```{r}


D_predicted %>% 
  ggplot(aes(x=genes_coding, y=n16, col=best_pred)) +
  geom_point() 
```

Join data again

```{r}
D_with_pred <- D_predicted %>% 
  select(species, best_pred) %>% 
  right_join(D, by="species")

# add sporeforming pred to normal sporeforming
D_final <- D_with_pred %>% 
  mutate(sporeforming = ifelse(is.na(sporeforming), best_pred, sporeforming))
           
  
```

Plot it finally (two different ways)

```{r}
D_final %>% 
  filter(phylum %in% c("Bacillota","Actinomycetota")) %>% 
  filter(!is.na(sporeforming)) %>% 
  ggplot(aes(x=genes_coding, y=n16, col=sporeforming)) +
  geom_point(size=1) +
  scale_color_manual(values = c("lightblue", "lightpink","blue","red")) +
  facet_wrap(~phylum)

D_final %>% 
  filter(phylum %in% c("Bacillota","Actinomycetota")) %>% 
  filter(!is.na(sporeforming)) %>% 
  mutate(sf_is_predicted = ifelse(sporeforming == "p1"|sporeforming =="p2", "yes","no")) %>% 
  mutate(sporeforming = ifelse(sporeforming == "1" | sporeforming == "p1", "1", 0)) %>% 
  ggplot(aes(x=genes_coding, y=n16, col=sporeforming, shape = sf_is_predicted)) +
  geom_jitter(size=1, height = 0.1) +
  facet_wrap(~phylum)
```

Inspect some on google.. It seems to be working :)) Both spore and non
spore :)

```{r}
D_final %>% 
  filter(sporeforming == "p1")

D_final %>% 
  filter(sporeforming == "p0")
```

Looking at missing data now

```{r}

p1 <- D_final %>% 
  select(phylum,sporeforming, oxygen.tolerance, growth_temp) %>% 
  filter(phylum %in% c("Bacillota","Actinomycetota")) %>% 
   vis_miss(cluster = T, sort_miss = T) +
    theme(
        plot.margin = margin(, 1.5, , , "cm")
  ) + ggtitle("after")
  
p2 <- D %>% 
  select(phylum,sporeforming, oxygen.tolerance, growth_temp) %>% 
    filter(phylum %in% c("Bacillota","Actinomycetota")) %>% 
   vis_miss(cluster = T, sort_miss = T) +
    theme(
        plot.margin = margin(, 1.5, , , "cm")
  ) + ggtitle("before")

p2 + p1
```

## Lastly we can set the gram negative phylums to non sporeforming

We only bother with those with over 10 seq
Lastly only bacillota and actinomycetota have sporeforming entries.. it seems in general
But just to make i checked the other entries

| phylum | has some sporeforming | source |
|--------|-----------------------|--------|

|"Bacillota" (firmicutes) | yes | <https://journals.asm.org/doi/full/10.1128/jb.00079-22>|
|"Bacteroidota"| no|<https://ami-journals.onlinelibrary.wiley.com/doi/full/10.1111/1462-2920.16094>|
|"Pseudomonadota" | no|<https://onlinelibrary.wiley.com/doi/abs/10.1002/9781118960608.gbm01157.pub2>|
|"Mycoplasmatota"|no|<https://www.sciencedirect.com/science/article/pii/B978008047378950018X>|
|"Actinomycetota" | yes | <https://www.mdpi.com/2076-2607/10/7/1471>|
|"Acidobacteriota"| gram-|<https://www.mdpi.com/2073-4441/14/24/4043>| 
|"Planctomycetota"| no mention(seems not) | - | 
|"Thermotogota" | no |<https://www.sciencedirect.com/science/article/pii/S0723202018304934?casa_token=8LPSXIb2dcsAAAAA:9RK8gm8BXAmh4cr925wGqqzO0_hRYKyS5XBjaTQVcAF7HT7z2EQWJ1eG6O5DryJbSNSh_qWqww>|
|"Spirochaetota" |gram -|<https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijsem.0.005690>| 
|"Fusobacteriota" | gram - |
<https://www.sciencedirect.com/science/article/pii/S2352513421003719>|
|"Chlamydiota" |no mention | -| 
|"Chlorobiota" | no meniton | -|
|"Deinococcota" |no|<https://journals.asm.org/doi/full/10.1128/mSystems.00786-20>|

```{r}
# seing which have more than 10 entries
D %>% 
  summarise(n=n(), .by=phylum) %>% 
  filter(n > 50) %>% 
  pull(phylum) %>% unique

# Defining the sporeforming correctly
has_sporeforming <- c("Bacillota","Actinomycetota")


# setting the others to not sporeforming
# They are all set to different values to differentiate where they came from eg, 1, 0, na0, p1 and p0 for statistics
D_final_many <- D_final %>% 
  mutate(sporeforming = case_when(
    phylum %in% has_sporeforming ~ sporeforming,
    sporeforming != NA ~ "20",
    .default = "na0"
  ))

D_final_many %>% 
  filter(phylum %in% has_sporeforming) %>% 
  ggplot(aes(phylum, fill = sporeforming)) + 
  geom_bar() 
```
The missing are either due to the model having a low chance of being correct, or due to entries having missing values.

```{r}
# Now cluster down the predicted to just 0 and 
D_to_print <- D_final_many %>% 
  mutate(sporeforming = case_when(
    sporeforming == "p0" ~ "0",
    sporeforming == "p1" ~ "1",
    sporeforming == "na0" ~ "0",
    .default = sporeforming
  ))


```

#

Write it to file
and vizualise. Now it looks a TON better!
```{r}
D_to_print %>% 
  select(oxygen.tolerance, growth_temp, sporeforming) %>% 
  vis_miss(cluster=T)

saveRDS(D_to_print, file = "./data/pred_bacdive_growth_ribdif.rds")
hist(D_to_print$ar_count)
```
